The Negative Effect of Traffic Noise on House Prices: A Landscape Hedonic Analysis

### 2014-07-02 reports/NoiseHedonicRaR.Rnw
nervous update

\usepackage[round]{natbib}
\usepackage{dcolumn}
\usepackage[nolists]{endfloat}
\usepackage[width = 5in]{geometry}
\usepackage{caption, amsmath, graphicx, setspace, multirow, color, hyperref, array}
\usepackage[modulo]{lineno}
%\renewcommand{\efloatseparator}{\mbox{}}

@@

\begin{figure}
\makebox[\textwidth][c]{\includegraphics[width = 1.2\textwidth]{../graphs/StudyAreaSaleValueDist}}
\caption{This figure shows the spatial extent of our study area as well as the significant spatial variation in single family house sales prices.}\label{fig:overview}
\end{figure}

@@

A map of our noise surface can be seen in Figure~??? Once the noise surface for the entire region was developed, we extracted the noise surface for the present study area by overlaying the housing parcel boundaries and calculating the mean noise level within the parcel.

@@

\subsection{Time}
We have sales transactions across six years (from 2005 to 2010) for our study area. Such temporal variation will allow us to estimate the hedonic price function over time and test for differences before, during, and after the economic recession. Table \ref{tab:SumStatsTime} displays the mean variable values by year of sale to look for differences over time. The mean sale price in 2009 (roughly \$230,000) is almost 20 percent lower than the mean sale price in 2006 (roughly \$280,000) and the number of sales transactions in 2009-10 is less than the number of transactions in 2005. However, most other variables have small or no differences across time. Additionally, univariate density plots of the variables over time also reveal no differences in the shape or spread of the distributions over time.\footnote{See Appendix A} For instance, it is not the case that after the recession the distribution ``hollowed out'' in the middle and became bimodal with many high- and low-priced houses selling with few mid-priced homes selling. Additionally, maps of the locations of sales transactions over time reveal no discernable patterns, either. The geographic dispersion of sales in earlier years look surprisingly similar to those in later years. Thus, we can find no evidence of only certain types or locations of houses transacting in later years compared to earlier years in our data.

@@

[Map of sales prices over time about here]

@@

We first conduct a Monte Carlo simulation recommended in \citet{Fotheringham2002} to see if the improved model fit with our LWR model was due to random chance. The intuition of the test is as follows - if the improved model fit isn't significant, then we should be able to see similar improvements in model fit by using the LWR model with the same houses but in a different spatial arrangement. Failing to recreate a spatial arrangement that yields a similar model fit as to that obtained with our actual data is then evidence that our spatial arrangement and the improved model fit obtained from a spatially heterogeneous model is significant.

@@

\subsection{Does the Implicit Price of Traffic Noise Change over Space?}
In the previous simulation we found that the hedonic regression function seems to exhibit spatial non-stationarity. However, we have not yet established which variables have spatially varying coefficients. In the second simulation, we seek to explore which of the hedonic regression coefficients exhibit spatial non-stationarity.

The mechanics of the second simulation are very similar to the those described in the previous section. We will randomly change the spatial arrangment of our house sales and estimate the LWR model. However, this simulation we will restrict the model to small bandwidths and compare the spread of the regression coefficient estimates under the random arrangment of the data to the spread found in our original LWR models with the actual spatial arrangement. If an independent variable in our hedonic regression function exhibits a stationary relationship with sales prices, then we should see similar levels of dispersion in the randomly arranged data as in our true data. However, if the dispersion we see in our random data is \emph{smaller} than the spread we observed in our actual data, that is evidence consistent with the variable having a non-stationary relationship. That is, the variation in regression coefficients obtained in our data is higher than we would observe from random chance.

@@

[Insert maps showing the spatial variation in the most important regression coefficients]

\subsection{Does the Implicit Price of Traffic Noise Change over Time?}

@@

In this section we look for patterns in our noise coefficients over time. We find evidence consistent with the estimated noise coefficients being more negative in the later time periods (housing bust) of our study than in the earlier periods (housing boom). This finding is robust to multiple different formulations and specifications. Table~\ref{tab:noisebetatime} presents the results of regressions using our estimated LWR noise coefficient as the dependent variable and time and spatial characteristics as explanatory variables.

Column (1) finds evidence of a negative linear trend in the LWR noise coefficients over time - predicting a LWR noise coefficient of approximately twice the magnitude for the end of our study period as for the beginning. Column (2) includes a quadratic term and again predicts noise coefficients much more negative towards the end of our study period compared to the beginning. We repeat the regressions from (1) and (2) in (3) and (4) but this time also include city fixed effects in order to control for any potential locational differences of house transactions over time. Again, we see very similar results - the noise coefficient is more negative in later time periods. We also limit our analysis to just the city of St.\ Paul and repeat the analysis from (1) and (2) in columns (5) and (6). Again, we estimate the noise coefficient to be more negative in later time periods.

% Requires LaTeX packages: dcolumn 
\begin{table}[!htbp] \centering 
  \caption{LWR Noise Coefficient Over Time} 
  \label{tab:noisebetatime} 
\scriptsize 
\begin{tabular}{@{\extracolsep{-1pt}}lD{.}{.}{6} D{.}{.}{7} D{.}{.}{7} D{.}{.}{7} D{.}{.}{7} D{.}{.}{7} } 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{6}{c}{\textit{Dependent variable: LWR Noise Coefficient}} \\ 
\cline{2-7} 
\\[-1.8ex] & \multicolumn{1}{c}{(1)} & \multicolumn{1}{c}{(2)} & \multicolumn{1}{c}{(3)} & \multicolumn{1}{c}{(4)} & \multicolumn{1}{c}{(5)} & \multicolumn{1}{c}{(6)}\\ 
\hline \\[-1.8ex] 
 Months since Jan 2005 & -0.00005^{***} & 0.00004^{***} & -0.00005^{***} & 0.00004^{***} & -0.0001^{***} & 0.0001^{***} \\ 
  & (0.000001) & (0.000003) & (0.000001) & (0.000003) & (0.000001) & (0.00001) \\ 
  & & & & & & \\ 
 (Months since Jan 2005)$^2$ &  & -0.000001^{***} &  & -0.000001^{***} &  & -0.000001^{***} \\ 
  &  & (0.00000004) &  & (0.00000004) &  & (0.0000001) \\ 
  & & & & & & \\ 
 Constant & -0.0016^{***} & -0.0029^{***} & -0.0021^{***} & -0.0034^{***} & -0.0016^{***} & -0.0035^{***} \\ 
  & (0.00003) & (0.0001) & (0.00003) & (0.0001) & (0.0001) & (0.0001) \\ 
  & & & & & & \\ 
\hline \\[-1.8ex] 
City Fixed Effects & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{No} \\ 
\hline \\[-1.8ex] 
Observations & \multicolumn{1}{c}{31,737} & \multicolumn{1}{c}{31,737} & \multicolumn{1}{c}{31,737} & \multicolumn{1}{c}{31,737} & \multicolumn{1}{c}{8,003} & \multicolumn{1}{c}{8,003} \\ 
R$^{2}$ & \multicolumn{1}{c}{0.1566} & \multicolumn{1}{c}{0.1722} & \multicolumn{1}{c}{0.3298} & \multicolumn{1}{c}{0.3446} & \multicolumn{1}{c}{0.1700} & \multicolumn{1}{c}{0.1915} \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{6}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
\end{tabular} 
\end{table}

@@

First and foremost, we need to be cautious about claims of causality. In an ideal research design, we would observe the same houses over time and randomly ``treat'' some houses with more or less noise to observe the change in price relative to the control houses. Such a study is not feasible and we are therefore limited to a hedonic analysis with observational data. Our conclusions, then, should be interpretted like most other hedonic analysis results and while we find a strong negative correlation between noise and house prices, stronger in some areas of our study area than others and stronger in some time periods than others, we cannot definitely state that noise causes lower sales prices.

Our work also faces data limitations. Notably, while our house sales data are collected over the course of six years, the traffic noise estimates are for the year 2007. To the extent that traffic flows and composition significantly changed over time, our traffic noise variables may be inaccurate for those other time periods. However, Table VM2 from the US. Department of Transportation Office of Highway Policy Information website reports that in the year 2007 there were approximately 30 million vehicle miles traveled in the urban areas of the state of MN, while for years 2008, 2009, and 2010 the respective values were 32.5, 32.3, and 32.0. Overall, this represents a change of less than 10 percent. \footnote{\url{http://www.fhwa.dot.gov/policyinformation/quickfinddata/qftravel.cfm}} Such small changes in total vehicle miles traveled are unlikely to yield significant differences in noise levels. Additionally, we do not believe the changes are likely to manifest in significant spatial variation in traffic noise, which is ultimately responsible for our identification of the noise regression coefficient. However, given enough time for changes in driving habits and technological advancements to gain significant market penetration (for instance, quieter hybrid cars), future work should seek to create time-series noise data in order to obtain even better estimates of the impact of traffic noise over time. Given the computational complexity of re-estimating the landscape noise surface, such work is beyond the scope of this paper.

>>>>

\usepackage[round]{natbib}
\usepackage{dcolumn}
\usepackage{rotating}
\usepackage{graphicx}
\usepackage[nolists]{endfloat}
\usepackage[width = 5in]{geometry}
\usepackage{caption, amsmath, graphicx, setspace, multirow, color, hyperref, array}
\usepackage[modulo]{lineno}
%\renewcommand{\efloatseparator}{\mbox{}}

@@

\begin{figure}
\makebox[\textwidth][c]{\includegraphics[width = 1.2\textwidth]{../graphs/Overview2column}}
\caption{This figure shows the spatial extent of our study area, the significant spatial variation in single family house sales prices, as well as our noise data.}\label{fig:overview}
\end{figure}

@@

A map of our noise surface can be seen in Figure~\ref{fig:overview}. Once the noise surface for the entire region was developed, we extracted the noise surface for the present study area by overlaying the housing parcel boundaries and calculating the mean noise level within the parcel.

@@

\subsection{Time}
We have sales transactions across six years (from 2005 to 2010) for our study area. Such temporal variation will allow us to estimate the hedonic price function over time and test for differences before, during, and after the economic recession. Table \ref{tab:SumStatsTime} displays the mean variable values by year of sale to look for differences over time. The mean sale price in 2009 (roughly \$230,000) is almost 20 percent lower than the mean sale price in 2006 (roughly \$280,000) and the number of sales transactions in 2009-10 is less than the number of transactions in 2005. However, most other variables have small or no differences across time.

@@

We also created univariate density plots of the variables over time to check for differences in the shape or spread of the distributions over time (see Figure~\ref{fig:DensityPlots}). It is not the case that after the recession the distribution ``hollowed out'' in the middle and became bimodal with many high- and low-priced houses selling with few mid-priced homes selling. Additionally, maps of the locations of sales transactions over time reveal no discernable patterns, either (see Figure~\ref{fig:SalesYear}). The geographic dispersion of sales in earlier years looks surprisingly similar to that of later years. Thus, we can find no evidence of only certain types or locations of houses transacting in later years compared to earlier years in our data.

\begin{figure}
\makebox[\textwidth][c]{\includegraphics[width = \textwidth]{../graphs/DensityPlotsByYear}}
\caption{This figure shows }\label{fig:DensityPlots}
\end{figure}

\begin{figure}
\makebox[\textwidth][c]{\includegraphics[width = \textwidth]{../graphs/SaleValue_with05}}
\caption{This figure shows }\label{fig:SalesYear}
\end{figure}

@@

We first conducted a Monte Carlo simulation recommended in \citet{Fotheringham2002} to see if the improved model fit with our LWR model was due to random chance. The intuition of the test is as follows - if the improved model fit isn't significant, then we should be able to see similar improvements in model fit by using the LWR model with the same houses but in a different spatial arrangement. Failing to recreate a spatial arrangement that yields a similar model fit as to that obtained with our actual data is then evidence that our spatial arrangement and the improved model fit obtained from a spatially heterogeneous model is significant.

@@

\subsection{Which Coefficients Vary over Space?}

The previous section presented evidence in support of the claim that there exists significant spatial variation in the hedonic function. However, we have not yet established which variables have spatially varying coefficients. Figures~\ref{fig:AcreBetas} -- \ref{fig:NoiseBetas} show the spatial variation in the estimated LWR coefficients for three important variables, lot size, house size, and noise. For comparison, we present the results across four different levels of local analysis, ranging from LWR bandwidths of 500 nearest houses to the nearest 2,000 sales.

A few patterns are clear in Figures~\ref{fig:AcreBetas} -- \ref{fig:NoiseBetas}. For instance, across all four bandwidths we see that the hedonic\footnote{Remember that the dependent variable is the log of sales price.} coefficient on lot size (measured in acres) is largest near the central business district of St.\ Paul. Thus, consistent with standard urban economic theory, to the extent that the hedonic coefficient is correlated with the price of land, our results suggest the price of land is highest near the central business district and declines with distance from the CBD. This tendency is consistent across the four different bandwidths presented (other specific results available upon request). The LWR coefficients on house size (measured in thousands of square feet of finished space) also seem to exhibit spatial variation across the different bandwidths. In the west-central portion of our study area (a collection of notably desirable neighborhoods) we see a cluster of large LWR coefficients, while those areas in the southernmost portion of our region (suburban areas) have the smallest coefficients.

In Figure~\ref{fig:NoiseBetas} we map the estimated LWR noise coefficients from four different bandwidths. In this figure the red areas represent large negative coefficients, while yellow areas are characterized by values closer to zero. It is interesting to note that for the smaller bandwidths, the area with the most negative noise coefficients is ringed by some of the coefficients closest to zero (and even greater than zero in some cases). As the bandwidth increases to 1000 or 2000 nearest houses, the variation in noise coefficients decreases substantially. In contrast to the previous two figures, we have largely eliminated the ``hot spots'' that appear in the smaller bandwidths.

\begin{figure}
 \makebox[\textwidth][c]{\includegraphics[width=1.1\textwidth]{../graphs/AcreBeta_allBandwidth}}
 \caption{Simula.}\label{fig:AcreBetas}
\end{figure}

\begin{figure}
 \makebox[\textwidth][c]{\includegraphics[width=1.1\textwidth]{../graphs/FinSqFtBeta_allBandwidth}}
 \caption{Simula.}\label{fig:HouseBetas}
\end{figure}

\begin{figure}
 \makebox[\textwidth][c]{\includegraphics[width=1.1\textwidth]{../graphs/NoiseBeta_allBandwidth}}
 \caption{Simula.}\label{fig:NoiseBetas}
\end{figure}

We now conduct a second series of simulation experiments aimed at testing whether the spatial variation exhibited in particular coefficients is significant or due to random chance. The mechanics of the second simulation are very similar to the those described in the previous section. We randomly change the spatial arrangment of our house sales and estimate the LWR model. However, in this simulation we restrict the model to small bandwidths and compare the spread of the regression coefficient estimates under the random arrangment of the data to the spread found in our original LWR models with the actual spatial arrangement. If an independent variable in our hedonic regression function exhibits a stationary relationship with sales prices, then we should see similar levels of dispersion in the randomly arranged data as in our true data. However, if the dispersion we see in our random data is \emph{smaller} than the spread we observed in our actual data, that is evidence consistent with the variable having a non-stationary relationship. That is, the variation in regression coefficients obtained in our data is higher than we would observe from random chance.

@@

\subsection{Does the Noise Coefficient Change over Time?}

@@

Figures~\ref{fig:AcreTime}--\ref{fig:NoiseTime} show the spatial distribution of some our LWR hedonic coefficients for bandwidths of 500 and 2,000 nearest sales within the last year for the sale years 2006--2010. Regardless of bandwidth, the figures show that the estimated LWR coefficients are different over time. In particular, the lot size and house size coefficients tend to decrease over time. What is surprising, however, is that the negative noise coefficient estimates tend to be more negative in later years of our study. Figure~\ref{fig:NoiseTime} shows that the spatial patterns are similar across years (the most negative coefficients tend to be in the same locations across years), but the average noise coefficient tends to be more negative for the later years.

\begin{sidewaysfigure}
 \includegraphics[trim = 0cm 2cm 0cm 0cm, clip = true, width = \textwidth]{../graphs/Acre_50_200_ByYear}
 \caption{Simula.}
 \label{fig:AcreTime}
\end{sidewaysfigure}

\begin{sidewaysfigure}
 \includegraphics[trim = 0cm 2cm 0cm 0cm, clip = true, width = \textwidth]{../graphs/FinFt_50_200_ByYear}
 \caption{Simula.}
 \label{fig:HouseTime}
\end{sidewaysfigure}

\begin{sidewaysfigure}
 \includegraphics[trim = 0cm 2cm 0cm 0cm, clip = true, width = \textwidth]{../graphs/Noise_50_200_ByYear}
 \caption{Simula.}
 \label{fig:NoiseTime}
\end{sidewaysfigure}

We also find statistical evidence consistent with the estimated noise coefficients being more negative in the later time periods. This finding is robust to multiple different formulations and specifications. Tables~\ref{tab:noisebetatime500} and \ref{tab:noisebetatime2000} present the results of regressions using our estimated LWR noise coefficient as the dependent variable and time and spatial characteristics as explanatory variables for LWR bandwidths of 500 and 2,000 nearest sales.

Both tables are structured similarly. Column (1) finds evidence of a negative linear trend in the LWR noise coefficients over time - predicting a LWR noise coefficient of approximately twice the magnitude for the end of our study period as for the beginning. Column (2) includes a quadratic term and again predicts noise coefficients much more negative towards the end of our study period compared to the beginning. We repeat the regressions from (1) and (2) in (3) and (4) but this time also include city fixed effects in order to control for any potential locational differences of house transactions over time. Again, we see very similar results - the noise coefficient is more negative in later time periods. We also limit our analysis to just the city of St.\ Paul and repeat the analysis from (1) and (2) in columns (5) and (6). Again, we estimate the noise coefficient to be more negative in later time periods.

% Requires LaTeX packages: dcolumn 
\begin{table}[!htbp] \centering 
  \label{tab:noisebetatime500} 
\scriptsize 
\begin{tabular}{@{\extracolsep{-1pt}}lD{.}{.}{6} D{.}{.}{7} D{.}{.}{7} D{.}{.}{7} D{.}{.}{7} D{.}{.}{7} } 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{6}{c}{\textit{Dependent variable: LWR Noise Coefficients (bandwidth = 500 nearest sales)}} \\ 
\cline{2-7} 
\\[-1.8ex] & \multicolumn{1}{c}{(1)} & \multicolumn{1}{c}{(2)} & \multicolumn{1}{c}{(3)} & \multicolumn{1}{c}{(4)} & \multicolumn{1}{c}{(5)} & \multicolumn{1}{c}{(6)}\\ 
\hline \\[-1.8ex] 
Months since Jan 2005 & -0.00005^{***} & 0.00003^{***} & -0.00005^{***} & 0.00003^{***} & -0.0001^{***} & 0.00004^{***} \\ 
  & (0.000001) & (0.000004) & (0.000001) & (0.000004) & (0.000002) & (0.00001) \\ 
  & & & & & & \\ 
 (Months since Jan 2005)$^2$ &  & -0.000001^{***} &  & -0.000001^{***} &  & -0.000001^{***} \\ 
  &  & (0.00000005) &  & (0.00000004) &  & (0.0000001) \\ 
  & & & & & & \\ 
 Constant & -0.0016^{***} & -0.0028^{***} & -0.0021^{***} & -0.0033^{***} & -0.0015^{***} & -0.0032^{***} \\ 
  & (0.00003) & (0.0001) & (0.00003) & (0.0001) & (0.0001) & (0.0002) \\ 
  & & & & & & \\ 
\hline \\[-1.8ex] 
City Fixed Effects & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{No} \\ 
\hline \\[-1.8ex] 
Observations & \multicolumn{1}{c}{31,737} & \multicolumn{1}{c}{31,737} & \multicolumn{1}{c}{31,737} & \multicolumn{1}{c}{31,737} & \multicolumn{1}{c}{8,003} & \multicolumn{1}{c}{8,003} \\ 
R$^{2}$ & \multicolumn{1}{c}{0.1273} & \multicolumn{1}{c}{0.1377} & \multicolumn{1}{c}{0.2931} & \multicolumn{1}{c}{0.3028} & \multicolumn{1}{c}{0.1566} & \multicolumn{1}{c}{0.1704} \\ 
\hline 
\hline \\[-1.8ex]  
\end{tabular} 
  \caption{LWR Noise Coefficient Over Time} 
\end{table}

% Requires LaTeX packages: dcolumn 
\begin{sidewaystable}[!htbp] \centering 
  \caption{LWR Noise Coefficient Over Time} 
  \label{tab:noisebetatime2000} 
\scriptsize 
\begin{tabular}{@{\extracolsep{-1pt}}lD{.}{.}{6} D{.}{.}{7} D{.}{.}{7} D{.}{.}{7} D{.}{.}{7} D{.}{.}{7} } 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{6}{c}{\textit{LWR bandwidth = 2,000 nearest sales}} \\ 
\cline{2-7} 
\\[-1.8ex] & \multicolumn{1}{c}{(1)} & \multicolumn{1}{c}{(2)} & \multicolumn{1}{c}{(3)} & \multicolumn{1}{c}{(4)} & \multicolumn{1}{c}{(5)} & \multicolumn{1}{c}{(6)}\\ 
\hline \\[-1.8ex] 
Months since Jan 2005 & -0.00005^{***} & 0.00003^{***} & -0.00005^{***} & 0.00003^{***} & -0.00005^{***} & 0.0001^{***} \\ 
  & (0.0000003) & (0.000002) & (0.0000003) & (0.000002) & (0.000001) & (0.000004) \\ 
  & & & & & & \\ 
 (Months since Jan 2005)$^2$ &  & -0.000001^{***} &  & -0.000001^{***} &  & -0.000001^{***} \\ 
  &  & (0.00000002) &  & (0.00000002) &  & (0.0000001) \\ 
  & & & & & & \\ 
 Constant & -0.0014^{***} & -0.0026^{***} & -0.0016^{***} & -0.0028^{***} & -0.0016^{***} & -0.0033^{***} \\ 
  & (0.00001) & (0.00003) & (0.00002) & (0.00003) & (0.00003) & (0.0001) \\ 
  & & & & & & \\ 
\hline \\[-1.8ex] 
City Fixed Effects & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{No} \\ 
\hline \\[-1.8ex] 
Observations & \multicolumn{1}{c}{31,737} & \multicolumn{1}{c}{31,737} & \multicolumn{1}{c}{31,737} & \multicolumn{1}{c}{31,737} & \multicolumn{1}{c}{8,003} & \multicolumn{1}{c}{8,003} \\ 
R$^{2}$ & \multicolumn{1}{c}{0.3828} & \multicolumn{1}{c}{0.4142} & \multicolumn{1}{c}{0.5128} & \multicolumn{1}{c}{0.5435} & \multicolumn{1}{c}{0.3454} & \multicolumn{1}{c}{0.3988} \\ 
\hline 
\hline \\[-1.8ex] 
\end{tabular} 
\end{sidewaystable}

@@

In a stronger research design, we would observe the same houses over time and randomly ``treat'' some houses with more or less noise to observe the change in price relative to the control houses. Such a study is not feasible and we are therefore limited to a hedonic analysis with observational data. Our conclusions, then, should be interpretted like most other hedonic analysis results and, while we find a strong negative correlation between noise and house prices, we cannot definitely state that noise causes lower sales prices.

Future work may want to further investigate this correlation for more evidence consistent with causation. For instance, by collecting data on the quality of house windows one could test for differential effects of more soundproof windows in noisier areas vs. quieter areas. Unfortunately, such house characteristics are unavailable for this study.

Our work faces other data limitations. Notably, while our house sales data are collected over the course of six years, the traffic noise estimates are for the year 2007. To the extent that traffic flows and composition significantly changed over time, our traffic noise variables may be inaccurate for those other time periods. However, Table VM2 from the US. Department of Transportation Office of Highway Policy Information website reports that in the year 2007 there were approximately 30 million vehicle miles traveled in the urban areas of the state of MN, while for years 2008, 2009, and 2010 the respective values were 32.5, 32.3, and 32.0. Overall, this represents a change of less than 10 percent. \footnote{\url{http://www.fhwa.dot.gov/policyinformation/quickfinddata/qftravel.cfm}} Such small changes in total vehicle miles traveled are unlikely to yield significant differences in noise levels. Additionally, we do not believe the changes are likely to manifest in significant spatial variation in traffic noise, which is ultimately responsible for our identification of the noise regression coefficient. However, given enough time for changes in driving habits and technological advancements to gain significant market penetration (for instance, quieter hybrid cars), future work should seek to create time-series noise data in order to obtain even better estimates of the impact of traffic noise over time. Given the computational complexity of re-estimating the landscape noise surface, such work is beyond the scope of this paper, but may be a worthwhile investment for future work seeking to test the robustness of our results.

### 2014-07-02 reports/NoiseHedonicRaR.tex
nervous update

First and foremost, we need to be cautious about claims of causality. In an ideal research design, we would observe the same houses over time and randomly ``treat'' some houses with more or less noise to observe the change in price relative to the control houses. Such a study is not feasible and we are therefore limited to a hedonic analysis with observational data. Our conclusions, then, should be interpretted like most other hedonic analysis results and while we find a strong negative correlation between noise and house prices, stronger in some areas of our study area than others and stronger in some time periods than others, we cannot definitely state that noise causes lower sales prices.

Our work also faces some serious data limitations. Notably, while our house sales data are collected over the course of six years, the traffic noise estimates are for the year 2007. To the extent that traffic flows and composition significantly changed over time, our traffic noise variables may be inaccurate for those other time periods. However, Table VM2 from the US. Department of Transportation Office of Highway Policy Information website reports that in the year 2007 there were approximately 30 million vehicle miles traveled in the urban areas of the state of MN, while for years 2008, 2009, and 2010 the respective values were 32.5, 32.3, and 32.0. Overall, this represents a change of less than 10 percent. \footnote{\url{http://www.fhwa.dot.gov/policyinformation/quickfinddata/qftravel.cfm}} Such small changes in total vehicle miles traveled are unlikely to yield significant differences in noise levels. Additionally, we do not believe the changes are likely to manifest in significant spatial variation in traffic noise, which is ultimately responsible for our identification of the noise regression coefficient. However, given enough time for changes in driving habits and technological advancements to gain significant market penetration (for instance, quieter hybrid cars), future work should seek to create time-series noise data in order to obtain even better estimates of the impact of traffic noise over time. Given the computational complexity of re-estimating the landscape noise surface, such work is beyond the scope of this paper.

>>>>

In a stronger research design, we would observe the same houses over time and randomly ``treat'' some houses with more or less noise to observe the change in price relative to the control houses. Such a study is not feasible and we are therefore limited to a hedonic analysis with observational data. Our conclusions, then, should be interpretted like most other hedonic analysis results and, while we find a strong negative correlation between noise and house prices, we cannot definitely state that noise causes lower sales prices.

Future work may want to further investigate this correlation for more evidence consistent with causation. For instance, by collecting data on the quality of house windows one could test for differential effects of more soundproof windows in noisier areas vs. quieter areas. Unfortunately, such house characteristics are unavailable for this study.

Our work faces other data limitations. Notably, while our house sales data are collected over the course of six years, the traffic noise estimates are for the year 2007. To the extent that traffic flows and composition significantly changed over time, our traffic noise variables may be inaccurate for those other time periods. However, Table VM2 from the US. Department of Transportation Office of Highway Policy Information website reports that in the year 2007 there were approximately 30 million vehicle miles traveled in the urban areas of the state of MN, while for years 2008, 2009, and 2010 the respective values were 32.5, 32.3, and 32.0. Overall, this represents a change of less than 10 percent. \footnote{\url{http://www.fhwa.dot.gov/policyinformation/quickfinddata/qftravel.cfm}} Such small changes in total vehicle miles traveled are unlikely to yield significant differences in noise levels. Additionally, we do not believe the changes are likely to manifest in significant spatial variation in traffic noise, which is ultimately responsible for our identification of the noise regression coefficient. However, given enough time for changes in driving habits and technological advancements to gain significant market penetration (for instance, quieter hybrid cars), future work should seek to create time-series noise data in order to obtain even better estimates of the impact of traffic noise over time. Given the computational complexity of re-estimating the landscape noise surface, such work is beyond the scope of this paper, but may be a worthwhile investment for future work seeking to test the robustness of our results.
